---
title: "Caterpillars"
author: "Marcus Becker"
date: "December 16, 2019"
output: html_document
---

```{r setup, include=FALSE}

# Load packages

library(DBI)
library(dbplyr)
library(dplyr)
library(dbplot)
library(ggplot2)
library(modeldb)
library(tidypredict)
library(config)
library(odbc)
library(keyring)
library(here)

# Load source functions (for now)
source(here("./R/connections.R"))

knitr::opts_chunk$set(echo = TRUE)

```

First step is to connect to the Wildtrax database. We'll use secured credentials via the `keyring` package, plus the `wt_con` function.

```{r}

wildtrax_con <- wt_con(db = "wildtrax", 
                    user = key_get("user", keyring = "wildtrax"),
                    password = key_get("password", keyring = "wildtrax"))

src_dbi(wt_con_2)

```

Now let's explore the Wildtrax database

```{r}

# List tables
dbListTables(wildtrax_con) # 87 tables
# List tables associated with a specific schema - e.g. aru schema
dbGetQuery(wildtrax_con, "SELECT table_name FROM information_schema.tables WHERE table_schema='aru'")
# List fields associated with a table
dbListFields(wildtrax_con, c("aru", "species_individual")) 

```

Queries

```{r}

spec_individ <- dplyr::tbl(wildtrax_con, dbplyr::in_schema("aru", "species_individual"))
recording <- dplyr::tbl(wildtrax_con, dbplyr::in_schema("aru", "recording"))
task <- dplyr::tbl(wildtrax_con, dbplyr::in_schema("aru", "task"))

# these are database objects -> not the physical download of the data (yet!)

yera_test <- spec_individ %>%
  inner_join(task, by = c("si_task_id" = "task_id")) %>%
  inner_join(recording, by = c("task_recording_id" = "recording_id")) %>%
  filter(si_species_code == "YERA") %>%
  # Retrieve the data for use in R
  collect()

```

Ideas for functions:

+ generic wildtrax connection function.
+ get_widltrax_data() - take a connection, build joins using dplyr syntax, does the joins, and return a tibble that's a pointer to the database. Tibble can be used from there.

```{r}

```

Disconnect from the Wildtrax database

```{r}

DBI::dbDisconnect(wt_con)

```





